---
title: "Processing MODIS Data"
author: "Beth E. Gerstner"
description: This script is part of the overarching Frugivoria project, which aims
  to build a comprehensive database of functional traits for birds and mammals (frugivores)
  in Central and South American tropical forests. This smaller project focuses on
  birds and mammals in Ecuador and will eventually be used to understand plant/animal
  interactions. Specifically, this script takes MODIS % forest cover data for four different tiles containing Ecuador and converts them from HDF files to a tif. It then stitches each tile together to create a single mosaicked tif for easy analyzing. This code is based off of https://www.hakimabdi.com/blog/working-with-hdf-eos-files-in-r-convert-project-and-mosaic. To choose appropriate tiles for study region, consult the MODIS instructions I wrote here
  :https://docs.google.com/document/d/1sDgXI9k5Dgo0rNtyRFj2HKhwwOq1B9Rc_jXo8qQr8PI/edit
date: "8/11/2020"
output: html_document
---

```{r}

library(gdalUtils)

# Function to make the conversion from HDF to Tiff
projHDF2GTiff = function(loc, hdfs, gtiffs, lyr, fromSRS, toSRS){ 
  if("gdalUtils" %in% rownames(installed.packages()) == FALSE){
    install.packages("gdalUtils", repos="http://r-forge.r-project.org")
    require(gdalUtils)
  } # install and load the gdalUtils package. 
  setwd(loc) # set the working directory to where the data is
  suppressWarnings(dir.create(paste(loc,"Projected",sep="/"))) # create a directory to store projected files
  for (i in 1:length(hdfs)){ 
    gdal_translate(hdfs[i],gtiffs[i],sd_index=lyr) # extract the specified HDF layer and save it as a Geotiff
    gdalwarp(gtiffs[i],paste(loc,"Projected",gtiffs[i],sep="/"),s_srs=fromSRS,t_srs=toSRS,srcnodata=-3000,dstnodata=-3000,overwrite = T) # project geotiffs
    unlink(gtiffs[i]) # delete unprojected geotiffs to save space
  }
}

```

# Use function to make the conversion with the data

```{r}
setwd("/Users/bethgerstner/Desktop/MODIS_2019") # where the individual MODIS tiles of interest are saved
myloc = "/Users/bethgerstner/Desktop/MODIS_2019/" # working directory
hdfs1 = list.files(getwd(), pattern="hdf$") # HDF file list
gtiffs1 = gsub("hdf","tif",hdfs1) # out out GeoTIFF file list
frm.srs = "+proj=sinu +lon_0=0 +x_0=0 +y_0=0 +a=6371007.181 +b=6371007.181 +units=m +no_defs" # original HDF SRS
to.srs = "+proj=longlat +datum=WGS84 +no_defs" # desired GeoTIFF SRS
# lyr is the HDF layer you want to extract. In this example it is "1" to 
# signify the first layer in the HDF file i.e. NDVI
# execute the function
# 
get_subdatasets("/Users/bethgerstner/Desktop/MODIS_2019/MOD44B.A2019065.h10v09.006.2020098203755.hdf") #to figure out what file you need to specify for lyr. Percent tree cover is layer 1.
projHDF2GTiff(loc = myloc, hdfs = hdfs1, gtiffs = gtiffs1, lyr = 1, fromSRS = frm.srs, toSRS = to.srs)
#rm(myloc,hdfs1,gtiffs1,frm.srs,to.srs,s.nodata,d.nodata) # remove variables to save memory
```

# Stitch tiles together to create mosaic

```{r}
mosaicGTiffs = function(proj.loc, gtiffs, mosaicName, overwrite){ 
  if("gdalUtils" %in% rownames(installed.packages()) == FALSE){ # checks if gdalutils is installed 
    install.packages("gdalUtils", repos="http://r-forge.r-project.org")
    require(gdalUtils)
  }
  suppressWarnings(dir.create(paste(proj.loc,"Mosaicked",sep="/"))) # creates a directory to store mosaicked file
  gdalwarp(gtiffs, paste(proj.loc,"/","Mosaicked","/",mosaicName,".tif",sep=""),overwrite = overwrite)
}
```

# Uses MOSAIC function above to stitch together the individual converted tif files.
```{r}
proj.loc = "/Users/bethgerstner/Desktop/MODIS_2019/Projected" # set the location of your GeoTIFFS
gtiffs2 = list.files("/Users/bethgerstner/Desktop/MODIS_2019/Projected",pattern = ".tif", full.names = T) # list the files
myMosaic = "MODIS_per_forest_2019" # the name of the final mosaicked GeoTIFF
# execute
mosaicGTiffs(proj.loc = proj.loc, gtiffs = gtiffs2, mosaicName = myMosaic, overwrite = T)
rm(proj.loc,gtiffs2,myMosaic) # remove variables to save memory
