---
title: "Frugivoria: Ecuadorian Plant richness "
author: "Hazel Anderson"
collaborators: "Phoebe Zarnestke, Beth Gernster, Pat Bills"
data input: "List of plant species in Ecuador in .csv, Range maps of those plant species in SpatialPolygonDataFrame, all forementioned data in RData"
data output: "Range maps in sf, final plot as png"
project: "Frugivoria"
date: "7/27/2020"
describtion: This script takes BIEN species range maps and uses them to create a map of species richness of Ecuador plant species.
output:
  pdf_document: default
  html_document: default
---

# Load required packages
```{r}
library(BIEN)
library(sp)
library(sf)
library(dplyr)
library(ggplot2)
library(scico)
library(rnaturalearth)
library(purrr)
library(ggsn)

```

# Set File Path
```{r}
data_path <- 'G:/Shared drives/SpaCE_Lab_FRUGIVORIA/r_code'
```


# Species Maps for All Ecuador Species
```{r}
if (file.exists("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/r_code/Ecuador_plant_range_maps_workspace.RData")){
    load("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/r_code/Ecuador_plant_range_maps_workspace.RData")
}else {
    #List of Ecuador Species
    Ecuador <- BIEN_occurrence_country("Ecuador")
    Ecuador_species_list <- unique(Ecuador$scrubbed_species_binomial)
    write.csv(Ecuador_species_list, "Ecuador_species_list.csv")
    
    #Species Maps for all plant species in Ecuador
    Ecuador_plant_species_list_vector <- unlist(Ecuador_species_list)
    Ecuador_plant_range_maps <- BIEN_ranges_load_species(species = Ecuador_plant_species_list_vector)
    
    # Converting the SpatialPolygonsdataframe to shapefiles
    Ecuador_plants <- st_as_sf(Ecuador_plant_range_maps)
    
    #Save workspace
    save.image("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/r_code/Ecuador_plant_range_maps_workspace.RData")
}
```

# Merge Ecuador plant species list with range map list
```{r}
mapped_list <- unlist(Ecuador_plant_range_maps$species)
#removing underscore from species names to match Ecuador_plant_species_list
mapped_list <- gsub("_", " ", mapped_list)

#making sure both lists are in the same form
mapped_list <- as_tibble(mapped_list)
Ecuador_species_list <- as_tibble(Ecuador_species_list)

#determining which species have not been mapped and saving list to csv
missing_maps <- setdiff(Ecuador_species_list, mapped_list)
write.csv(missing_maps, "plant_missing_maps.csv")
```

# Load in world map
```{r}
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")

# country subset. In this case we are removing the Galapagos by defining the bounding box around the Ecuador polygon.
CRpoly <- worldMap %>% filter(sovereignt == "Ecuador")
ecuador_shp <-st_crop(CRpoly, c(xmin=-84, xmax=-75.24961, ymin=-4.990625, ymax=1.455371))

# trim the map to appropriate study area. This zooms out of Ecuador, which will be easier for visualization down the line.
limsCR <- st_buffer(ecuador_shp, dist = 1) %>% st_bbox()

# countries
divpolPlot <-
ggplot() +
geom_sf(data = ecuador_shp) +
coord_sf(
xlim = c(limsCR["xmin"], limsCR["xmax"]),
ylim = c(limsCR["ymin"], limsCR["ymax"])
) +
scale_x_continuous(breaks = c(-84)) +
theme(
plot.background = element_rect(fill = "#ffffff"),
panel.background = element_rect(fill = "#ffffff"),
panel.grid = element_blank(),
line = element_blank(),
rect = element_blank()
)
divpolPlot
```

# Plot hulls
```{r}
hullsPlot <-
ggplot() +
geom_sf(data = ecuador_shp) +
geom_sf(data = Ecuador_plants, aes(fill = species), alpha = 0.7) +
scale_fill_scico_d(palette = "davos", direction = -1, end = 0.9, guide = FALSE) +
coord_sf(
xlim = c(limsCR["xmin"], limsCR["xmax"]),
ylim = c(limsCR["ymin"], limsCR["ymax"])
) +
scale_x_continuous(breaks = c(-84)) +
theme(
plot.background = element_rect(fill = "#FFFFFF"),
panel.background = element_rect(fill = "#FFFFFF"),
panel.grid = element_blank(),
line = element_blank(),
rect = element_blank()
)
hullsPlot

```

# Define a grid
```{r}
CRGrid <- ecuador_shp %>%
st_make_grid(cellsize = 0.2) %>%
st_intersection(ecuador_shp) %>%
st_cast("MULTIPOLYGON") %>%
st_sf() %>%
mutate(cellid = row_number())
```

# Richness for convex hulls
```{r}
richness_gridEOO <- CRGrid %>%
st_join(Ecuador_plants) %>%
mutate(overlap = ifelse(!is.na(id), 1, 0)) %>%
group_by(cellid) %>%
summarize(num_species = sum(overlap))
```

# Empty grid, more detailed in nature
```{r}
gridPlot <-
ggplot() +
geom_sf(data =ecuador_shp) +
geom_sf(data = CRGrid) +
coord_sf(
xlim = c(limsCR["xmin"], limsCR["xmax"]),
ylim = c(limsCR["ymin"], limsCR["ymax"])
) +
scale_x_continuous(breaks = c(-84)) +
theme(
plot.background = element_rect(fill = "#FFFFFF"),
panel.background = element_rect(fill = "#FFFFFF"),
panel.grid = element_blank(),
line = element_blank(),
rect = element_blank()
)
gridPlot
```

# Richness for convex hulls, more detailed
```{r}
gridRichCR_eoo <-
ggplot(richness_gridEOO) +
geom_sf(data = ecuador_shp, fill = "grey", size = 0.1) +
geom_sf(aes(fill = num_species), color = NA) +
scale_fill_scico(palette = "davos", direction = -1, end = 0.9) +
coord_sf(
xlim = c(limsCR["xmin"], limsCR["xmax"]),
ylim = c(limsCR["ymin"], limsCR["ymax"])
) +
scale_x_continuous(breaks = c(-84)) +
theme(
plot.background = element_rect(fill = "#FFFFFF"),
panel.background = element_rect(fill = "#FFFFFF"),
panel.grid = element_blank(),
line = element_blank(),
rect = element_blank(),
axis.title.x = element_blank(),
axis.title.y = element_blank()
) + 
scalebar(ecuador_shp, dist = 200, st.dist=.05, st.size=4, height=0.02, transform = TRUE, dist_unit = "km", model = 'WGS84') +
north(ecuador_shp, location="topright", scale=0.15, symbol=3) +
labs(fill = "richness") 

gridRichCR_eoo

```

# Saving richness map
```{r}
ggsave(filename = "Plant richness map Ecuador.png", dpi = 300, device="png")
```

#Investigate the overlap of species ranges with forest cover
```{r}
library(fasterize)
library(raster)
#Load in forest cover data for 2019
for_cov<- raster("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/analysis/MODIS_data_2019/MODIS_per_forest_2019.tif")
# Run fasterize to turn the richness map into a raster 
raster_grid <- fasterize(
  richness_gridEOO,
  for_cov,
  field = "num_species",
  fun = "sum",
  background = NA,
)
#Write the richness raster to a file 
writeRaster(raster_grid, filename="G:/Shared drives/SpaCE_Lab_FRUGIVORIA/analysis/MODIS_data_2019/plant_ric.tif", format="GTiff")
#read in raster
raster_grid <- raster("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/analysis/MODIS_data_2019/plant_ric.tif")
```

#To look at trends in forest cover over the species richness surface, they have to have the same resolution so we must resample the richness raster to make it match forest cover (finer res)
```{r, figures-side, fig.show="hold", out.width="50%"}
#Resample the diversity raster
raster_rich_fin_res<-resample(raster_grid, for_cov, method="bilinear")
#make sure extents match
fc_mask <- mask(for_cov, raster_rich_fin_res)
#remove water (value 200)
fc_mask[fc_mask==200] <-NA
#Subset the forest cover raster to areas <50 and >50% forest cover 
fc_less_50 <- fc_mask <= 50
#remove 0s to allow a mask where just the values of 1 are being used 
fc_less_50[fc_less_50 == 0] <- NA
fc_great_50 <-fc_mask > 50
#remove 0s to allow a mask where just the values of 1 are being used 
fc_great_50[fc_great_50 == 0] <- NA
plot(fc_less_50, main="forest cover <= 50%")
plot(fc_great_50, main="forest cover > 50%")
# Take richness layer and mask by these forest cover cover subsets to have richness found above and below these thresholds
rich_less_50 <-  mask(raster_rich_fin_res,fc_less_50)
rich_great_50 <- mask(raster_rich_fin_res, fc_great_50)
  
#Plots of taxonomic richness in forest/non-forest
plot(rich_less_50, main="taxonomic richness in areas of <=50% forest cover", add=T)
plot(rich_great_50, main="taxonomic richness in areas > 50% forest cover")
#Summary of richness over those two forest thresholds
summary(values(rich_less_50))
summary(values(rich_great_50))
hist((values(rich_less_50)), xlab="taxonomic richness", ylab="frequency", main= "Taxonomic richness in regions <= 50% forest cover")
hist((values(rich_great_50)), xlab="taxonomic richness", ylab="frequency", main= "Taxonomic richness in regions > 50% forest cover")
```
```