---
title: 'Frugivoria: Ecuadorian Birds, Mammals, and Plants'
author: "Beth Gerstner & Hazel Anderson"
date: "6/1/2020"
output:
  pdf_document: default
  html_document: default
---
# Load packages and IUCN API
```{r, message=FALSE}
library(httr) 
library(jsonlite)
library(dplyr)
library(rredlist) 

token <- '3b3db1c753a7cad0616e16146363ec88eead97d185cb3304070d7343123516fd'
```

# Pull in Ecuadorian species from IUCN Redlist
```{r}
#Ecuador
ecuador_species <- rl_sp_country("EC", key = token, parse = TRUE)
c1 <- ecuador_species$result[1]

```

# Break down by species group for the world (will subset to Ecuadorian species)
```{r}

# Get codes for the birds
all_birds <-rl_comp_groups(group = c('birds'), key = token)

# Get codes for the mammals
all_mammals <-rl_comp_groups(group = c('mammals'), key = token)

```

# Mammals and birds occurring in Ecuador 
```{r}

#Mammals

#South America
#extract IUCN results for all mammals and birds from list
all_mammals_df <- as.data.frame(all_mammals[['result']])
all_birds_df <- as.data.frame(all_birds[['result']])

#Mammals
ec_mammals <-all_mammals_df %>%
  filter(taxonid %in% c1$taxonid)


#Birds
ec_birds <-all_birds_df %>%
  filter(taxonid %in% c1$taxonid)

```

# Get habitat for each species --------------------------------------------
# want lowland montane and highland montane species

```{r}
# Include error-checking code in the function so that it will return NA for everything if there's no data.

get_habitat <- function(ID) {
    habitat_by_sp<- rl_habitats(id=ID,key=token)$result
    if (class(habitat_by_sp) == 'data.frame') {
        data.frame(taxonid=ID, habitat_by_sp)
    } else {
        data.frame(taxonid=ID)
    }
}
```

# Apply get_habitat to each row of ec_mammals
```{r}

ec_habitat_mammals <- ec_mammals %>%
  rowwise %>%
  do(get_habitat(.$taxonid))

```

# Apply get_habitat to each row of_birds

```{r}
ec_habitat_birds <- ec_birds %>%
  rowwise %>%
  do(get_habitat(.$taxonid))


```

# Merge the habitat data with the IUCN list so that we have the statuses and habitats combined
```{r}

mammal_IUCN_full<- merge.data.frame(ec_mammals, ec_habitat_mammals, by= "taxonid", all=TRUE)

bird_IUCN_full<- merge.data.frame(ec_birds, ec_habitat_birds, by= "taxonid", all=TRUE)


```

# Pull in Elton Traits data to begin subsetting to frugivores

```{r}
#Read in Elton Traits dataset 
birds <- read.csv("/Volumes/GoogleDrive/Shared drives/SpaCE_Lab_FRUGIVORIA/L0/Elton_traits/BirdFuncDat.csv")
mamm <- read.csv("/Volumes/GoogleDrive/Shared drives/SpaCE_Lab_FRUGIVORIA/L0/Elton_traits/MamFuncDat.csv")
```

# Combine bird elton traits dataset (full) and IUCN bird dataset for Ecuador together
```{r}

#Combine mammal elton traits dataset and IUCN mammal data together
colnames(mamm)[which(names(mamm) == "Scientific")] <- "scientific_name"
mamm_trait_IUCN <- merge.data.frame(mamm, mammal_IUCN_full, by= "scientific_name", all.y=TRUE)

#Combine bird elton traits dataset and IUCN bird data together
colnames(birds)[which(names(birds) == "Scientific")] <- "scientific_name"
bird_trait_IUCN<- merge.data.frame(birds, bird_IUCN_full, by= "scientific_name", all.y=TRUE)

```

# Go through the merges that did not work correctly (kept all IUCN species that were not found in the Elton Traits database) and compare against synonym lists

```{r}




```



#____________________________________________________________________________________________
#### Hazel start your code here! I'll continue adding above

# Load packages
```{r}
library(BIEN)
library(ape)
library(maps)
library(sp)
```
# Pull Ecuador plant species list from BIEN
```{r}
Ecuador <- BIEN_occurrence_country("Ecuador")
```

# Trait lists
```{r}
fruit_type <- BIEN_trait_trait(trait = "fruit type")
minimum_fruit_length <- BIEN_trait_trait(trait = "minimum fruit length")
maximum_fruit_lenth <- BIEN_trait_trait(trait = "maximum fruit length")
plant_fruiting_duration <- BIEN_trait_trait(trait = "plant fruiting duration")
```
