---
title: "Plant Trait L0 to L1"
author: "Hazel Anderson"
date: "6/11/2020"
output: html_document
---

# Load required packages
```{r}
library(BIEN)
library(ape)
library(maps)
library(sp)
```
# Pull Ecuador plant species list from BIEN
```{r}
Ecuador <- BIEN_occurrence_country("Ecuador")
```

#List of all avaliable traits in BIEN
```{r}
BIEN_trait_list()
```

# Create a vector of all the Trait Names.  The BIEN functions will take a vector of traits instead of just one


```{r}

trait_vector <- c( "diameter at breast height (1.3 m)", 
    "flower color", 
    "flower pollination syndrome", 
    "fruit type", 
    "inflorescence length", 
    "leaf area", 
    "leaf area per leaf dry mass", 
    "leaf carbon content per leaf dry mass", 
    "leaf carbon content per leaf nitrogen content", 
    "leaf compoundness", 
    "leaf dry mass", 
    "leaf dry mass per leaf fresh mass", 
    "leaf fresh mass", 
    "Leaf lamina fracture toughness", 
    "leaf life span", 
    "leaf nitrogen content per leaf area", 
    "leaf nitrogen content per leaf dry mass", 
    "leaf phosphorus content per leaf area", 
    "leaf phosphorus content per leaf dry mass", 
    "leaf photosynthetic rate per leaf area", 
    "leaf photosynthetic rate per leaf dry mass", 
    "leaf relative growth rate", 
    "leaf stomatal conductance for H2O per leaf area", 
    "leaf thickness", 
    "longest whole plant longevity", 
    "maximum fruit length", 
    "maximum leaf length", 
    "maximum leaf width", 
    "maximum whole plant height", 
    "maximum whole plant longevity", 
    "minimum fruit length", 
    "minimum leaf length", 
    "minimum leaf width", 
    "minimum whole plant height", 
    "plant flowering begin", 
    "plant flowering duration", 
    "plant fruiting duration", 
    "root dry mass", 
    "seed length", 
    "seed mass", 
    "stem dry mass", 
    "stem relative growth rate", 
    "stem wood density", 
    "vessel lumen area", 
    "vessel number", 
    "whole plant dispersal syndrome", 
    "whole plant growth form", 
    "whole plant growth form diversity", 
    "whole plant height", 
    "whole plant primary juvenile period length", 
    "whole plant sexual system", 
    "whole plant vegetative phenology", 
    "whole plant woodiness"
)
# diameter_at_breast_height <- BIEN_trait_trait(trait = "diameter at breast height (1.3 m)")
# flower_color <- BIEN_trait_trait( trait = "flower color")
# flower_pollination_syndrome <- BIEN_trait_trait(trait = "flower pollination syndrome")
# fruit_type <- BIEN_trait_trait(trait = "fruit type")
# inflorescence_length <- BIEN_trait_trait(trait = "	inflorescence length")
# leaf_area <- BIEN_trait_trait(trait = "leaf area")
# leaf_area_per_leaf_dry_mass <- BIEN_trait_trait(trait = "leaf area per leaf dry mass")
# leaf_carbon_content_per_leaf_dry_mass <- BIEN_trait_trait(trait = "leaf carbon content per leaf dry mass")
# leaf_carbon_content_per_leaf_nitrogen_content <- BIEN_trait_trait(trait = "leaf carbon content per leaf nitrogen content")
# leaf_compoundness <- BIEN_trait_trait(trait = "leaf compoundness")
# leaf_dry_mass <- BIEN_trait_trait(trait = "leaf dry mass")
# leaf_dry_mass_per_leaf_fresh_mass <- BIEN_trait_trait(trait = "leaf dry mass per leaf fresh mass")
# leaf_fresh_mass <- BIEN_trait_trait(trait = "leaf fresh mass")
# Leaf_lamina_fracture_toughness <- BIEN_trait_trait(trait = "Leaf lamina fracture toughness")
# leaf_life_span <- BIEN_trait_trait(trait = "leaf life span")
# leaf_nitrogen_content_per_leaf_area <- BIEN_trait_trait(trait = "leaf nitrogen content per leaf area")
# leaf_nitrogen_content_per_leaf_dry_mass <- BIEN_trait_trait(trait = "leaf nitrogen content per leaf dry mass")
# leaf_phosphorus_content_per_leaf_area <- BIEN_trait_trait(trait = "leaf phosphorus content per leaf area")
# leaf_phosphorus_content_per_leaf_dry_mass <- BIEN_trait_trait(trait = "leaf phosphorus content per leaf dry mass")
# leaf_photosynthetic_rate_per_leaf_area <- BIEN_trait_trait(trait = "leaf photosynthetic rate per leaf area")
# leaf_photosynthetic_rate_per_leaf_dry_mass <- BIEN_trait_trait(trait = "leaf photosynthetic rate per leaf dry mass")
# leaf_relative_growth_rate <- BIEN_trait_trait(trait = "leaf relative growth rate")
# leaf_stomatal_conductance_for_H2O_per_leaf_area <- BIEN_trait_trait(trait = "leaf stomatal conductance for H2O per leaf area")
# leaf_thickness <- BIEN_trait_trait(trait = "leaf thickness")
# longest_whole_plant_longevity <- BIEN_trait_trait(trait = "longest whole plant longevity")
# maximum_fruit_lenth <- BIEN_trait_trait(trait = "maximum fruit length")
# maximum_leaf_length <- BIEN_trait_trait(trait = "maximum leaf length")
# maximum_leaf_width <- BIEN_trait_trait(trait = "maximum leaf width")
# maximum_whole_plant_height <- BIEN_trait_trait(trait = "maximum whole plant height")
# maximum_whole_plant_longevity <- BIEN_trait_trait(trait = "maximum whole plant longevity")
# minimum_fruit_length <- BIEN_trait_trait(trait = "minimum fruit length")
# minimum_leaf_length <- BIEN_trait_trait(trait = "minimum leaf length")
# minimum_leaf_width <- BIEN_trait_trait(trait = "minimum leaf width")
# minimum_whole_plant_height <- BIEN_trait_trait(trait = "minimum whole plant height")
# plant_flowering_begin <- BIEN_trait_trait(trait = "plant flowering begin")
# plant_flowering_duration <-BIEN_trait_trait(trait = "plant flowering duration")
# plant_fruiting_duration <- BIEN_trait_trait(trait = "plant fruiting duration")
# root_dry_mass <- BIEN_trait_trait(trait = "root dry mass")
# seed_length <- BIEN_trait_trait(trait = "seed length")
# seed_mass <- BIEN_trait_trait(trait = "seed mass")
# stem_dry_mass <- BIEN_trait_trait(trait = "stem dry mass")
# stem_relative_growth_rate <- BIEN_trait_trait(trait = "stem relative growth rate")
# stem_wood_density <- BIEN_trait_trait(trait = "stem wood density")
# vessel_lumen_area <- BIEN_trait_trait(trait = "vessel lumen area")
# vessel_number <- BIEN_trait_trait(trait = "vessel number")
# whole_plant_dispersal_syndrome <- BIEN_trait_trait(trait = "whole plant dispersal syndrome")
# whole_plant_growth_form <- BIEN_trait_trait(trait = "whole plant growth form")
# whole_plant_growth_form_diversity <- BIEN_trait_trait(trait = "whole plant growth form diversity")
# whole_plant_height <- BIEN_trait_trait(trait = "whole plant height")
# whole_plant_primary_juvenile_period_length <- BIEN_trait_trait(trait = "whole plant primary juvenile period length")
# whole_plant_sexual_system <- BIEN_trait_trait(trait = "whole plant sexual system")
# whole_plant_vegetative_phenology <- BIEN_trait_trait(trait = "whole plant vegetative phenology")
# whole_plant_woodiness <- BIEN_trait_trait(trait = "whole plant woodiness")
```

## Unique species list for Ecuador

count of species in Ecuador list: 

```{r}
Ecuador_species_list <- unique(Ecuador$scrubbed_species_binomial)
print(length(Ecuador_species_list))
```


## Test geting a list of all traits for one species (first in Ecuaador list).  

```{r}
species <- Ecuador_species_list[1]
```


Traits for `r species` : 

```{r}
traits <- BIEN_trait_traitbyspecies(species, trait_vector)
```



## Traits for species in Ecuador

working on this...

```{r}
# for (i in Ecuador_species_list) {
#   BIEN_trait_species(species = "i")
# }
```

#Unique list of fruit types
```{r}
fruit_type <- BIEN_trait_trait(trait = "fruit type")
fruit_type_list <- unique(fruit_type$trait_value)
print(fruit_type_list)
```
